---
title: Skimming through {skimr}
author: Qiushi Yan
date: '2019-11-22'
slug: introduction-to-skimr

summary: 'Develpoed by the rOpenSci conmmunity, skimr is designed to provide useful, tidy summary statistics of data frames'

bibliography: ../bib/introduction-to-skimr.bib
link-citations: true
lastmod: '2020-04-05T23:27:44+08:00'

image:
  caption: 'Photo by [Josh Sorenson](https://www.pexels.com/@joshsorenson)'
  focal_point: ''
  preview_only: no

---


```{r, include = FALSE}
knitr::opts_chunk$set(message = F, warning = FALSE,
                      collapse = TRUE, cache = TRUE,
                      comment = "#>")
```



# `skim()` basics  {#intro}


The **skimr**[@R-skimr] package is developed by the [rOpenSci](https://github.com/ropensci) cummunity, with its core function `skim()` as an enhanced version of `summary()`, and has just experienced a major update to V2. For a more comprehensive guide towards its features, please visit its [vignette](https://docs.ropensci.org/skimr/index.html). ^[In the sense that `skim()` provides useful, tidy summary statistics and displays it in a pretty form for exploratory analysis based on data frames. `skim()` cannot supplant `summary()` in terms of model interpretation, viewing geospatial objects, etc.] 

What `skim()` returns can be divided into 2 parts:  

  * **Data summary**: an overview of the input data frame, including **#** of rows and columns, column types and if the data frame is grouped  
  * **Summary of column variables based on their types, with one section per variable type and one row per variable**: For instance, all numerical variables are explained by missing rate, quantiles (`fivenum()`), mean, standard deviation and a tiny inline histogram. With factors, the count of top values are returned instaed.  

```{r}
library(skimr)
library(tidyverse)
```

```{r}
skim(iris)
```

The second argument `...` in `skim()` could specify any number of variables to be included, similar to `dplyr::select()`: 

```{r}
skim(iris, Sepal.Width, Species)
```

or with common `select()` helpers:  

```{r}
skim(iris, contains("Length"))
```


We can also take the `summary()` of the skimmed data frame, and get the **Data Summary** part:   

```{r}
skim(iris) %>% 
  summary()
```


`skim()` also supports grouped data created by `dplyr::group_by()`(and later we will see that `skim()` workes seamlessly in a typical **tidyverse**[@tidyverse2019] workflow). In the grouped case, one additional column for each grouping variable is added.

```{r}
iris %>%
  group_by(Species) %>%
  skim()
```

To better illustrate `skim()`, we should know it returns a `skim_df` object, which is essentially **a single wide data frame** combining the results, with some additional attributes and two metadata columns:  

* skim_type: class of the variable
* skim_variable: name of the original variable  
  

`is_skim_df()` function is used to assert that an object is a skim_df.

```{r}
class(skim(iris))
is_skim_df(skim(iris))
```

To explicitly show 2 columns `skim_type` and `skim_variable` and see the nature of this `skim_df` object, we can transfrom it to a `tibble` or `data.frame`  

```{r}
iris %>% 
  skim() %>% 
  as_tibble()
```

As you can see, the **Data Summary** part are excluded after the transformation, and the wide data frame are left. This is in contrast to `summary.data.frame()`, which stores statistics in a table. The distinction is important, **because the `skim_df object` is pipeable and easy to use for additional manipulation**: for example, the user could select all of the `variable` means, or all summary statistics for a specific variable (Note that we don't have to really transform it before such manipulation).  

Filtering by column types:  

```{r}
skim(iris) %>% 
  filter(skim_type == "numeric")
```


```{r}
skim(iris) %>% 
  filter(skim_variable == "Sepal.Length")
```


When using `select()` to choose statistics we want to display, `n_missing` and `complete_rate` can be directly specified, since they are computed for all types of columns. Other statistics must be specified in `select` with a prefix, as shown in the column names of `skim(iris) %>% as_tibble()` (i.e, extracting the arithmet mean by `numeric.mean`):

```{r}
# no need to put a prefix
skim(iris) %>% 
  select(skim_variable, skim_type, n_missing, complete_rate)
```


```{r}
# extract the mean and median with a prefix
skim(iris) %>% 
  filter(skim_type == "numeric") %>%
  select(skim_variable, numeric.mean, numeric.p50)
# extract top levels of a factor
skim(iris) %>% 
  filter(skim_type == "factor") %>% 
  select(skim_variable, factor.top_counts)
```

Or we can combine `select_if()` with `skim()`  

```{r}
iris %>% 
  select_if(is.numeric) %>% 
  skim()
```




Finally, just for fun, what statistics or which variable we choose to use will not affect the `skim_df` variable type, yet this is not the case for `skim_variable` and `skim_type`, which are intrinsic to `skim_df` objects. If we exclude these two, the `skim_df` object will be automatically coerced to a tibble:

```{r}
skim(iris) %>% 
  select(-skim_type, -skim_variable) %>%
  is_skim_df()
```


# Manipulating the results  {#manipulation}

As noted above, `skim()` returns a wide data frame, an `skim_df` object. This is usually the most sensible format for the majority of operations when investigating data, and **skimr** also provides functions aimed at facilitating the exploratory process .  

First, `partition()` returns a named list of the wide data frames **for each data type**. Unlike the original data the partitioned data only has columns corresponding to the skimming functions use for this data type. These data frames are, therefore, not `skim_df` objects. And we can easily subset that list to get desirable component:  

```{r}
iris %>%
  skim() %>%
  partition()
```

Each component of the resulting list is an `one_skim_df` object, and can be readily coerced to tibbles:  

```{r}
iris %>%
  skim() %>%
  partition() %>%
  map(class)
```

This is useful when we only want to look at certain data types, we can subset that component out of the partitioned result, coerce it to a tibble, and then manipulate it. We don't have to put a prefix before column names now, since rows in every component are of identical data types, and therefore share same statistics:  

```{r}
# subset
iris_numeric <- iris %>%
  skim() %>% 
  partition() %>%
  pluck("numeric") %>%
  as_tibble()

iris_numeric %>% 
  arrange(sd)  ## no more numeric.sd
```

Second, `yank()` selects only specified variables base on their type, this can be viewed as a shortcut to `partition()` and subsetting:   

```{r}
iris %>% 
  skim() %>% 
  yank("factor")
```

Use `focus()` to select columns of the skimmed results and keep them as a `skim_df`; it always keeps the metadata column. In other words, to **focus** on some statistics:  

```{r}
iris %>% 
  skim() %>%
  focus(n_missing, numeric.mean)
```

# Customizing `skim()`  {#customization}

We can create out own skimming function with the `skim_with()` function factory. To add a statistic for a data type, create an `sfl()` (a skimr function list) for each class that you want to change:  

```{r}
# also look at skewness and kurtosis for numerical variables
library(e1071)
my_skim <- skim_with(numeric = sfl("3rd_central_moment" = skewness,
                                   "4th_central_moment" = kurtosis))
```

Now, we can use `my_skim` to skim through numerical variables:  

```{r}
my_skim(iris) %>% 
  yank("numeric")
```

When using `skim_with()`, the named list in `sf1()` correspond to statistics calculated on a specified type, so names of that list must fall into one of R's classes, available classes are (see `?get_skimmers`):  

* numeric
* character  
* factor  
* logical  
* complex  
* Date   
* POSIXct  
* difftime  
* ts  
* AsIs  

It you want to customize `skim()` to support more data classes, [here](https://docs.ropensci.org/skimr/articles/Supporting_additional_objects.html#conclusion) is an simple example, adding support for `sf` oject.  



# Skimming non-data frames  {#other}

In **skimr** [v2](https://github.com/ropensci/skimr/releases), `skim()` will attempt to coerce non-data frames (such as vectors and matrices) to data frames. In most cases with vectors, the object being evaluated should be equivalent to wrapping the object in as.data.frame().  


Skimming charactor vector:  
```{r}

skim(letters)
```

This is same to: 

```{r}
skim(as.data.frame(letters))
```

The way `skim()` works with `list` is some natural, for data frames are just a special case of list:  
```{r}
my_list <- list(a = 1:10, b = letters[1:10], c = factor(1:10))
skim(my_list)
```



skimming `ts` object: 

```{r}
lynx
# note the last inline line graph, pretty cute
skim(lynx)
```

when skimming matrices, columns and treated as variables and rows as observations, so statistics are calculated on a column basis:  

```{r}
m <- matrix(1:12, nrow = 4, ncol = 3)
m
```

```{r}
skim(m)
```

Transpose the matrix:  

```{r}
skim(t(m))
```

# References {#ref}
