---
title: 'Reordering within facets'
author: Qiushi Yan
date: '2019-11-29'
slug: reorder-varible-within-facets-reorder-within

image:
  caption: "Photo by [Leah Kelley](https://www.pexels.com/@leah-kelley-50725)"
  focal_point: ""
  preview_only: true
categories:
  - R

summary: '{ggetxt} provides convenient yet less-recognized graphing helpers to reorder an x or y axis within facets.'
lastmod: '2020-04-05T21:38:24+08:00'
bibliography: ../bib/reorder-within.bib
link-citations: true
---

This blog post is inspired by Julia Silge's [
Reordering and facetting for ggplot2
](https://juliasilge.com/blog/reorder-within/). By reproducing the original post using another different dataset, I hope this could still shed some light on a handful of ggplot2 tips.   



# Common words in Jane Austen's books  

```{r, include = FALSE}
knitr::opts_chunk$set(message = F, warning = TRUE, 
                      comment = "#>", collapse = TRUE)
```

Here is a simple task: we want to find the top-10 most common words in several Jane Austen's books, draw a bar graph, and then put them into different facets.  

```{r}
library(tidyverse)
library(tidytext)
library(janeaustenr)

theme_set(theme_light())
```




Here we use `unnest_tokens` from the `tidytext` package [@R-tidytext] to break the text into individual tokens and transform it into a tidy structure:  

```{r}
# calculate 10 most common words in each book
common_words <- austen_books() %>%
  unnest_tokens(word, text) %>% 
  anti_join(stop_words) %>% 
  count(book, word) %>% 
  group_by(book) %>% 
  top_n(10)

common_words
```

For details many and other useful text mining techniques in R, I recommend reading [Text mining with R: A tidy approach](Text Mining with R: A tidy approach) by Julia Silge and David Robinson[-@silge2017text].  

```{r}
# plot common words
common_words %>% 
  ggplot(aes(y = word, 
             x = n, 
             fill = book)) +
  geom_col(show.legend = FALSE) + 
  facet_wrap(~ book, ncol = 2, scales = "free_y") + 
  theme(axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 14)) +
  rcartocolor::scale_fill_carto_d(palette = "PurpOr")
```
(**Reminder**: don't forget `scales = "free_y"` or `scales = "free"` in this case.)

In its default setting, ggplot2 puts the names in alphabetical order, because they are of type character.  One common solution is to use `fct_reorder()` or its equivalent `reorder()` in baes R`:  

```{r}
common_words %>% 
  ggplot(aes(y = fct_reorder(word, n), 
             x = n, 
             fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ book, ncol = 2, scales = "free_y") + 
  theme(axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 14)) +
  rcartocolor::scale_fill_carto_d(palette = "PurpOr")
```

Well, our plot improve a bit, but is still wanting in some respects. In each facet, the order of words is not strictly in line with its frequency. In fact, `fct_reorder` or `reorder` only reorder all of these together, meaning that `word` are reordered in an overall level so that it cannot guarantee order in any specific subset of data.  

# Helper functions for reordering  

tidytext provides 2 helper functions to reorder a variable on a facet level.

`reorder_within()` takes three arguments:

- `x`: the item we want to reorder  
- `by`: what we want to reorder by  
- `within`: the groups or categories we want to reorder within  

After that, we use `scale_x_reordered()`(or `scale_y_reordered()`) to finish the plot, these 2 scales can take any argument in common scale functions(e.g., `name`, `labels`, `expand`).

Here is the final plot:  
```{r}
# reorder word by n, within each book
common_words %>% 
  ggplot(aes(y = reorder_within(word, n, book), 
             x = n, 
             fill = book)) +
  geom_col(show.legend = FALSE) + 
  scale_y_reordered() + 
  facet_wrap(~ book, ncol = 2, scales = "free_y") + 
  theme(axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 14)) +
  rcartocolor::scale_fill_carto_d(palette = "PurpOr") +
  labs(title = "What are the most common words in Jane Austen's books?",
       x = NULL,
       y = NULL)
```



Without `scale_y_reordered` which overides ggplot2's default scale, the label of the reordered variable will carry a suffix related to the group variable, making lables way too long and uninformative. Here is an simple example to demonstrate this problem using the [babynames](https://github.com/hadley/babynames) data: 

```{r}
library(babynames)
```


```{r}
top_names <- babynames %>% 
  filter(year >= 1950,
           year < 1990) %>% 
  mutate(decade = (year %/% 10) * 10) %>% 
  group_by(decade) %>%
  count(name, wt = n) %>% 
  group_by(decade) %>% 
  top_n(10)

top_names %>% 
  ggplot(aes(y = reorder_within(name, n, decade), 
             x = n, 
             fill = factor(decade))) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~ decade, scales = "free_y") +
  labs(x = NULL, y = NULL)
```




# References  

Nowosad, J. (2018). 'CARTOColors' Palettes. R package version 1.0.0. https://nowosad.github.io/rcartocolor