---
title: Text Classfication with Penalized Logistic Regression
author: Qiushi Yan
date: '2020-05-02'
slug: text-classification-logistic
categories: ["Data Analysis", "Machine Learning", "R"]
summary: 'Teach models how to distinguish Charlotte Brontë from Emily Brontë'
lastmod: '2020-05-02T23:33:49+08:00'
bibliography: ../bib/text-classification.bib
biblio-style: apalike
link-citations: yes
image:
  caption: ''
  focal_point: ''
  preview_only: yes
---


In this post I aim to train a text classification model with penalized logistic regression using the [`tidymodels`](https://www.tidymodels.org/) framework. Data are from 5 books and  downloaded via the [`gutenbergr`](https://docs.ropensci.org/gutenbergr/) package, written by either Emily Brontë or Charlotte Brontë. And the goal is to predict the author of a line, in other words the probability of line being written by one sister instead of another.  

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      comment = "#>",
                      collapse = TRUE,
                      cache = TRUE)
```


```{r, cache = FALSE}
library(tidyverse)
library(tidytext)
library(gutenbergr)
```



```{r, eval = FALSE}
books <- gutenberg_works() %>% 
  filter(str_detect(author, "Brontë, Emily|Brontë, Charlotte")) %>% 
  gutenberg_download(meta_fields = c("title", "author")) %>% 
  transmute(title,
            author = if_else(author == "Brontë, Emily", 
                             "Emily Brontë", 
                             "Charlotte Brontë") %>% factor(),
            line_index = row_number(),
            text)
```

```{r, cache = FALSE, echo = FALSE}
theme_bar <- function(base_size = 12,
                      strip_text_size = 14,
                      strip_text_margin = 5,
                      subtitle_size = 14,
                      subtitle_margin = 10,
                      plot_title_size = 18,
                      plot_title_margin = 10,
                      ...) {
  ret <- ggplot2::theme_minimal(base_size = base_size, ...)
  ret$strip.text <- ggplot2::element_text(
    hjust = 0, size = strip_text_size,
    margin = ggplot2::margin(b = strip_text_margin)
  )
  ret$plot.subtitle <- ggplot2::element_text(
    hjust = 0, size = subtitle_size,
    margin = ggplot2::margin(b = subtitle_margin)
  )
  ret$plot.title <- ggplot2::element_text(
    hjust = 0, size = plot_title_size,
    margin = ggplot2::margin(b = plot_title_margin)
  )
  ret
}

theme_set(theme_bar())
```

```{r, echo = FALSE}
books <- read_rds("D:/RProjects/data/blog/bronte-sisters-data.rds")
```

`books` is at line level

```{r}
books
```


To obtain tidy text structure illustrated in [Text Mining with R](https://www.tidytextmining.com/), I use `unnest_tokens()` to perform tokenization and remove all the stop words. I also removed characters like `'`, `'s`, `'` and whitespaces to return valid column names after widening. But it turns out this served as some sort of stemming too! (heathcliff's becomes heathcliff). Then low frequency words (whose frequency is less than 0.05% of an author's total word counts) are removed. The cutoff may be a little too high if you plot that histogram, but I really need this to save computation efforts on my laptop :sweat_smile:. 

```{r}
clean_books <- books %>% 
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>% 
  filter(!str_detect(word, "^\\d+$")) %>% 
  mutate(word = str_remove_all(word, "_|'s|'|\\s"))
  
total_words <- clean_books %>%
  count(author, name = "total")

tidy_books <- clean_books %>%
  left_join(total_words) %>% 
  group_by(author, total, word) %>%
  filter((n() / total) > 0.0005) %>% 
  ungroup() 
```

# Comparing word frequency

Before building an actual predictive model, let's do some EDA to see different tendency to use a particular word! This will also shed light on what we would expect from the text classification. Now, we will compare word frequency (proportion) between the two sisters.

```{r, fig.width = 15, fig.height = 9}
tidy_books %>% 
  group_by(author, total) %>%
  count(word) %>% 
  mutate(prop = n / total) %>% 
  ungroup() %>% 
  select(-total, -n) %>%
  pivot_wider(names_from  = author, values_from = prop,
              values_fill = list(prop = 0)) %>% 
  ggplot(aes(x = `Charlotte Brontë`, y = `Emily Brontë`, 
             color = abs(`Emily Brontë` -  `Charlotte Brontë`))) + 
  geom_jitter(width = 0.001, height = 0.001, alpha = 0.2, size = 2.5) + 
  geom_abline(color = "gray40", lty = 2) + 
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5, size = 7.5) + 
  scale_color_gradient(low = "darkslategray4", high = "gray75") +
  scale_x_continuous(labels = scales::label_percent()) + 
  scale_y_continuous(labels = scales::label_percent()) +  
  theme(legend.position = "none") + 
  coord_cartesian(xlim = c(0, NA)) + 
  labs(title = "Word frequency between two sisters") + 
  theme(text = element_text(size = 18),
        plot.title.position = "plot")
```


Words lie near the line such as "home", "head" and "half" indicate similar tendency to use that word, while those that are far from the line are words that are found more in one set of texts than another, for example "headthcliff", "linton", "catherine", etc. 

What does this plot tell us? Judged only by word frequency, it looks that there are a number of words that are quite characteristic of Emily Brontë (upper left corner). Charlotte, on the other hand, has few representative words (bottom right corner). We will investigate this further in the model. 

# Modeling

## Data preprocessing  

There are `r nrow(dplyr::count(tidy_books, word))` and features (words) and `r nrow(dplyr::count(tidy_books, line_index))` observations in total. Approximately 18% of the response are 1 (Emily Brontë).

```{r}
tidy_books %>% 
  count(author) %>% 
  mutate(prop = n / sum(n))
```


Now it's time to widen our data to reach an appropriate model structure, this similar to a document-term matrix, with rows being a line and column word count. 

```{r}
library(tidymodels)
set.seed(2020)
doParallel::registerDoParallel()

model_df <- tidy_books %>% 
  count(line_index, word) %>% 
  pivot_wider(names_from = word, values_from = n,
              values_fill = list(n = 0)) %>% 
  left_join(books, by = c("line_index" = "line_index")) %>% 
  select(-title, -text)

model_df
```



## Train a penalized logistic regression model 

Split the data into training set and testing set.

```{r}
book_split <- initial_split(model_df)
book_train <- training(book_split)
book_test <- testing(book_split)
```

Specify a L1 penalized logistic model, center and scale all predictors and combine them in to a `workflow` object.  

```{r}
logistic_spec <- logistic_reg(penalty = 0.05, mixture = 1) %>%
  set_engine("glmnet")

book_rec <- recipe(author ~ ., data = book_train) %>% 
  update_role(line_index, new_role = "ID") %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())

book_wf <- workflow() %>% 
  add_model(logistic_spec) %>% 
  add_recipe(book_rec)

initial_fit <- book_wf %>% 
  fit(data = book_train)
```


`initial_fit` is a simple fitted regression model without any hyperparameters. By default `glmnet` calls for 100 values of lambda even if I specify $\lambda = 0.05$. So the extracted result aren't that helpful.

```{r}
initial_fit %>%
  pull_workflow_fit() %>% 
  tidy()
```

We can make predictions with `initial_fit` anyway, and examine metrics like overall accuracy. 

```{r}
initial_predict <- predict(initial_fit, book_test) %>% 
    bind_cols(predict(initial_fit, book_test, type = "prob")) %>%
    bind_cols(book_test %>% select(author, line_index))

initial_predict
```


How good is our initial model?

```{r}
initial_predict %>% 
  accuracy(truth = author, estimate = .pred_class)
```

Nearly 84% of all predictions are right. This isn't a very statisfactory result since "Charlotte Brontë" accounts for 81% of `author`, making our model only slightly better than a classifier that would assngin all `author` with "Charlotte Brontë" anyway. 

### Tuning lambda

We can figure out an appropriate penalty using resampling and tune the model. 

```{r}
logistic_wf_tune <- book_wf %>%
  update_model(logistic_spec %>% set_args(penalty = tune()))

lambda_grid <- grid_regular(penalty(), levels = 100)
book_folds <- vfold_cv(book_train, v = 10)
```

Here I build a set of 10 cross validations resamples, and set `levels = 100` to try 100 choices of lambda ranging from 0 to 1.  

Then I tune the grid: 
```{r}
logistic_results <- logistic_wf_tune %>%    
  tune_grid(
    resamples = book_folds,
    grid = lambda_grid)
```

There is an `autoplot()` method for the tuned results, but I instead plot two metrics versus lambda respectivcely by myself. 
```{r, fig.height = 10, fig.width = 10}
logistic_results %>% 
  collect_metrics() %>% 
  mutate(lower_bound = mean - std_err,
         upper_bound = mean + std_err) %>%
  ggplot(aes(penalty, mean)) + 
  geom_line(aes(color = .metric), size = 1.5, show.legend = FALSE) + 
  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound)) + 
  facet_wrap(~ .metric, nrow = 2) + 
  labs(y = NULL,
       x = expression(lambda),
       title = "Performance metric of logistic regession versus differenct choices of L1 regularization")
```

Ok, the two metrics both display a monotone decrease as lambda increases, but does not exhibit much change once lambda is greater than 0.1, which is essentailly random guess according to the author's respective proportion of appearance in the data. This plot shows that the model is generally better at small penalty, suggesting that the majority of the predictors are fairly important to the model. We may lean towards larger penalty with slightly worse performance, bacause they lead to simpler models. It follows that we may want to choose lambda in top rows in the following data frame
```{r}
top_models <- logistic_results %>% 
    show_best("roc_auc", n = 100) %>%
    arrange(desc(penalty)) %>% 
    filter(mean > 0.9)

top_models
```

`select_best()` with return the 9th row with $\lambda \approx 0.000586$ for its highest performance on `roc_auc`. But I'll stick to the parsimonious principle and pick $\lambda \approx 0.00376$ at the cost of a fall in roc_auc by 0.005 and in accuracy by 0.001. 

```{r}
logistic_results %>% 
  select_best(metric = "roc_auc")

book_wf_final <- finalize_workflow(logistic_wf_tune,
                                  parameters = top_models %>% slice(1))
```


Now the model specification in the workflow is filled with the picked lambda: 
```{r, echo = TRUE}
book_wf_final %>% pull_workflow_spec()
```

The next thing is to fit the best model with the training set, and evaluate against the test set. 




```{r, echo = TRUE}
logistic_final <- last_fit(book_wf_final, split = book_split)

logistic_final %>% 
  collect_metrics()

logistic_final %>% 
  collect_predictions() %>%
  roc_curve(truth = author, `.pred_Emily Brontë`) %>% 
  autoplot()
```

The accuracy of our logisitc model rises by a rough 7% to 93.8%, with `roc_auc` being nearly 0.904. This is pretty good!

There is also the confusion matrix to check. The model does well in identifying Charlotte Brontë (low false positive rate, high sensitivity), yet suffers relatively high false negative rate (mistakenly identify 39% of Emily Brontë as Charlotte Brontë, low specificity). 
```{r, echo = TRUE}
logistic_final %>%
  collect_predictions() %>%
  conf_mat(truth = author, estimate = .pred_class) 
```


To evaluate predictors, I agian use `fit` and `pull_workflow` to extract model fit. Variable importance plots implemented in the [vip](https://koalaverse.github.io/vip/index.html) package provides an intuitive way to visualize importance of predictors in this scenario, using th absolute value of the t-statistic as a measure of VI.


```{r, fig.height = 11, fig.width = 11, fig.cap = "Variable importance plot for penalized logistic regression"}
library(vip)

logistic_vi <- book_wf_final %>% 
  fit(book_train) %>% 
  pull_workflow_fit() %>%
  vi(lambda = top_models[1, ]$penalty) %>% 
  group_by(Sign) %>% 
  top_n(30, wt = abs(Importance)) %>%
  ungroup() %>% 
  mutate(Sign = if_else(Sign == "POS", 
                        "More Emily Brontë", "More Charlotte Brontë"))

logistic_vi %>% 
  ggplot(aes(y = reorder_within(Variable, abs(Importance), Sign),
             x = Importance)) + 
  geom_col(aes(fill = Sign), show.legend = FALSE) +
  scale_y_reordered() + 
  facet_wrap(~ Sign, nrow = 1, scales = "free") + 
  labs(title = "How word usage classifies Brontë sisters",
       x = NULL,
       y = NULL) + 
  theme(axis.text = element_text(size = 18),
        plot.title = element_text(size = 24),
        plot.title.position = "plot")
```

Is it cheating to use names of a character to classify authors? Perhaps I should consider include more books and remove names for text classification next time.  

Note that variale importance in the left panel is generally smaller than the right, this corresponds to what we find in the [word frequency](#comparing-word-frequency) plot that Emily Brontë has more and stronger characteristic words.

Anyway, the model with the best lambda seems quite powerful in distinguishing these two authors. I look forward to build a multinomial classification model in `tidymodels` to include Anne Brontë some other time!



