---
title: Analyzing COVID-19 Publications
author: Qiushi Yan
date: '2020-04-06'
slug: [analyzing-covid19-publications]
summary: 'A simple text analysis on the abstract of up-to-date COVID-19 publications.'
lastmod: '2020-05-08T12:54:24+08:00'
featured: no
bibliography: ../bib/covid19.bib
link-citations: yes

categories:
  - Data Analysis
  - R

image:
  caption: 'Image by [cottonbro](https://www.pexels.com/@cottonbro)'
  focal_point: ''
  preview_only: no

---


In this post, I will be performing a simple text analysis on the abstract of publications on the coronavirus disease (COVID-19), courtesy of [WHO](https://www.who.int/emergencies/diseases/novel-coronavirus-2019/global-research-on-novel-coronavirus-2019-ncov). We begin by steps of data preprocessing.  


```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE,
                      comment = "#>", 
                      collapse = TRUE)
```


# Data preprocessing

```{r}
library(tidyverse)
```

```{r, echo = FALSE}
theme_bar <- function(base_size = 11,
                      strip_text_size = 12,
                      strip_text_margin = 5,
                      subtitle_size = 13,
                      subtitle_margin = 10,
                      plot_title_size = 16,
                      plot_title_margin = 10,
                      ...) {
  ret <- ggplot2::theme_minimal(base_size = base_size, ...)
  ret$strip.text <- ggplot2::element_text(
    hjust = 0, size = strip_text_size,
    margin = ggplot2::margin(b = strip_text_margin)
  )
  ret$plot.subtitle <- ggplot2::element_text(
    hjust = 0, size = subtitle_size,
    margin = ggplot2::margin(b = subtitle_margin)
  )
  ret$plot.title <- ggplot2::element_text(
    hjust = 0, size = plot_title_size,
    margin = ggplot2::margin(b = plot_title_margin)
  )
  ret
}

theme_set(theme_bar())
```


```{r}
raw <- read_csv("D:/RProjects/data/covid-research.csv") %>% 
  janitor::clean_names() 

glimpse(raw)
```


For simplicity I ignore many of the vairables (mostly for identification) and rows with missing values on `abstract`. I was a little disappointied to find out that `published_month` are all missing, otherwise we may see a trend of some sort on research topics there. One remaining problem is that some of the papers are not written in English. 
The `detect_language()` from  the [`cdl3`](https://github.com/ropensci/cld3) can detect language forms at a fairly high success rate. It's a R wrapper for Google's Compact Language Detector 3, which is a neural network model for language identification.  

```{r}
library(cld3)
raw %>% 
  count(language = detect_language(abstract), sort = TRUE)
```

In this post I keep only the rows that were classified as "en". Also, as illuatrated in [Text Mining with R](https://www.tidytextmining.com), text analysis commonly requires preprocessing steps like tokenizing, eliminating stop words and word stemming. I added custom keywords and did some maunual transformation to make up for misclassifications by `detect_language`, but there will still be non-English words, though. My transformations (say, both "covid" and "19" now become "covid19") will certainly induce errors, but there is no better workaround I could think of now.

```{r}
library(tidytext)
library(SnowballC)

words <- raw %>% 
  filter(detect_language(abstract) == "en") %>% 
  unnest_tokens(word, abstract) %>% 
  mutate(word = wordStem(word)) %>%
  mutate(word = case_when(
    word == "19" ~ "covid19",
    word == "covid" ~ "covid19",
    word == "coronaviru" ~ "coronavirus",
    word == "viru" ~ "virus",
    word == "epidem" ~ "epidemic",
    word == "studi" ~ "study",
    word == "respiratori" ~ "respiratory",
    word == "emetin" ~ "emetine",
    word == "acut" ~ "acute",
    word == "sever" ~ "severe",
    word == "manag" ~ "manage",
    word == "hospit" ~ "hospital",
    word == "diseas" ~ "disease",
    word == "deceas" ~ "disease",
    word == "caus" ~ "cause",
    word == "emerg" ~ "emerge",
    word == "includ" ~ "include", 
    word == "dai" ~ "wet nurse",
    word == "ncovid" ~ "ncov",
    word == "countri" ~ "country",
    word == "provid" ~ "provide",
    word == "peopl" ~ "people",
    TRUE ~ word
  )) %>% 
  anti_join(stop_words %>% 
            add_row(word = c( "2", "1",  "dub", "thi", "ha", "wa", "检查", "cd", "gt",
                              "lt", "tnt", "thei"), 
                    lexicon = "custom")) %>% 
  filter(!(str_detect(word, "^\\d+$") | str_detect(word, "^\\d+\\w$")))


words
```


# Common words and keywords extraction  

An immediate question is, what are the most common words among all these publications?  

```{r, fig.height = 18, fig.width = 12}
words %>% 
  count(word, sort = TRUE) %>%
  top_n(50) %>%
  ggplot(aes(y = fct_reorder(word, n),
             x = n)) + 
  geom_col() + 
  scale_x_continuous(expand = c(0.01, 0)) + 
  labs(y = NULL,
       x = "# of words",
       title = "Top 50 common words in COVID-19 publications") +
  theme(text = element_text(size = 18),
        plot.title.position = "plot",
        plot.title = element_text(size = 35, face = "bold"),
        axis.ticks.y = element_blank())
```

I'm also interested in paper-specific properties, namely their keywords, what topics distinguish them from others? In comparison to the commonly used algorithm tf-idf, I prefer using weighted log odds proposed by @monroe_colaresi_quinn, which a standardized metric from a complete statistical model. It is also implemented in the R package [`tidylo`](https://github.com/juliasilge/tidylo)[@tidylo]. The reason is that tf-idf cannot extract the varying use trend of common words, if a word appears in every research paper, then its inverse document frequency will be zero. For weighted log odds this is not the case, even if all researched mentioned this word it can still differentiate those who used it a lot more often from those who used less. This could be essential when we are trying to find an emphasis on which researchers place as our understanding of the virus advances. Sadly I have no access to the exact date of the publication, so I will just display words with topest score and their corresponding publications.

```{r}
library(tidylo)

words %>%
  count(title, word) %>% 
  bind_log_odds(set = title, feature = word, n = n) %>%
  top_n(20) %>% 
  knitr::kable()
```

Many words listed here are acronyms in biology, chemistry and medicine. For example, "hljby" is a type of porcine epidemic diarrhoea virus which "pedv" stands "for", fpv means Favipiravir (a type of drug), and "tlr" represents Toll-like receptors. 

# Fit a LDA topic model  

Let's then fit a 5-topic LDA topic model, before that we should convert the data frame to a docuemnt term matrix using `cast_dtm`. There are various implementations of this kind of model, here I use `stm::stm`. The choice of $K$ (number of topics) here is somewhat arbitrary here, but [Julia Silge](https://juliasilge.com/) had a great [post](https://juliasilge.com/blog/evaluating-stm/) about it.    

```{r}
dfm <- cast_dfm(words %>% count(title, word),
                term = word,
                document = title,
                value = n)
```

`dfm` is a document-term matrix with 2415 documents and 13383 features.  

```{r, eval = FALSE}
library(stm)
topic_model <- stm(dfm, K = 5, init.type = "LDA", verbose = FALSE)
```

```{r, echo = FALSE}
topic_model <- pins::pin_get("covid-topic-model")
```


Topic-term probability distributions are accessed by `tidy()`, this gives a glance of the underlying meaning of these topics:
```{r, fig.height = 12, fig.width = 12}
# topic-term distribution
tidy(topic_model) %>% 
  group_by(topic) %>% 
  top_n(10) %>% 
  ungroup() %>%
  mutate(topic = factor(topic) %>% str_c("topic", .)) %>% 
  ggplot(aes(y = reorder_within(term, beta, topic),
         x = beta,
         fill = topic)) + 
  geom_col(show.legend = FALSE) + 
  scale_y_reordered() + 
  facet_wrap(~ topic, scales = "free_y", nrow = 3) + 
  labs(y = NULL,
       x = "Docuemtn-term probabilities",
       title = "A 6-topic LDA model") + 
  theme(text = element_text(size = 18),
        plot.title = element_text(size = 30, face = "bold"),
        strip.text = element_text(size = 25, hjust = 0.05),
        axis.ticks.y = element_blank())
```

It's hard to interpret these topics without domain knwoledge. But it seems to me that topic3 is related to clinical findings, topic4 to china and wuhan, the epicenter of covid19.

# A network of paired words  

Another question of interest is the relationship between words: what group of words tend to appear together? I look at the [phi coefficient](https://en.wikipedia.org/wiki/Phi_coefficient), which is essentailly $\chi^2$ statistc in a contingency table applied to categorical variables.  

As each abstract is a natual unit of measure, a pair of words that both appear in the same abstract are seen as "appearing together". We could compute $\phi$ based on pairwise counts:

```{r}
library(widyr)

word_cors <- words %>% 
  add_count(word) %>% 
  filter(n > 20) %>%
  select(-n) %>%
  pairwise_cor(item = word, feature = title, sort = TRUE)

word_cors
```

A network visualization of word correlation is a good idea:  

```{r, fig.width = 12, fig.height = 12}
library(ggraph)
library(tidygraph)

word_cors %>% 
  filter(correlation > 0.4) %>% 
  as_tbl_graph() %>% 
  ggraph(layout = "fr") + 
  geom_edge_link(aes(alpha = correlation), show.legend = FALSE) + 
  geom_node_point(color = "lightblue", size = 6.5) + 
  geom_node_text(aes(label = name), repel = TRUE, size = 5.5)
```

As you can see, there are still some non-English words that stemming and adding stopwrods cannot handle... Nonetheless, we are be able to identify some of the clusters revovling around infant infection (infant, pregnant, newborn, mother), pathology (angiotensin, protein, receptor), symptoms (lung, thicken, lesion), etc.  

# References