---
title: Who is Carrying There Team in OWL?
author: Qiushi Yan
date: '2020-04-25'
categories: ["Data Analysis", "Overwatch", "R"]
summary: "Fleta deadlifts in all 3 positions"
lastmod: '2020-04-25T00:09:45+08:00'
featured: no
bibliography: ../bib/fleta-deadlift.bib
biblio-style: apalike
link-citations: yes
image:
  focal_point: ''
  preview_only: yes
draft: yes
---

# The Fleta deadlift

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE,
                      comment = "#>", 
                      collapse = TRUE,
                      cache = TRUE)
```

The [Overwatch Stats Lab](https://overwatchleague.com/en-us/statslab) is a treasure trove holding vast digital riches for people who want to observe the Overwatch League in a more detailed manner. One of my favourite statistic is the **Fleta deadlift**, calculated as

$$
\text{Fleta deadlifts} = \frac{\text{final blows of a player}}{\text{final blows of the whole team}}
$$
It comes as no surprise that the index was named after Fleta, the current Shanghai dragons DPS player, who is known for his former amazing performance in a less talented team before he joined Seoul Dyanasty and entered OWL. If you are not familiar with esports or overwatch, I have some equivelent "Fleta deadlifted his team" examples in the NBA context: 

- Michael Jordan in the 1998 final  
- Kobe Bryant (I miss him) in season 2006-2007, or perhaps half time in his career  
- Lebron Jmaes in the 2015 final  


The player section on Overwatch Stats Lab even established a special tab for this stat, showing records with Fleta deadlift > 50% (see the last tab below):  
```{r, echo = FALSE}
knitr::include_url("https://overwatchleague.com/en-us/statslab-players")
```

It should be obvious though, that Fleta deadlifts are more suitable for describe performaces of DPS players, and less applicable to the other two positions, tanks and supports. Since final blows may not be a good measure for a player's performance if his main responsibility indicates otherwise. 


I used `load_data.R` to laod in data and do soem preprocessing. You can browse it at. 
```{r}
library(tidyverse)
library(vroom)

df <- vroom("D:/RProjects/data/overwatch/fleta_deadlifts.csv")
```

## Extend the stat to tanks and supports  


However, since OWL mandates the 2-2-2 composition only after the third stage in season 2019. A player can belong to different positions even in the same map or match.
```{r}
dpss <- c("Hanzo", "Junkrat", "Tracer", "Soldier: 76", "Widowmaker", "McCree", "Pharah", "Genji", "Sombra", "Reaper", "Doomfist", "Bastion", "Mei", "Torbjörn", "Symmetra", "Ashe")
tanks <- c("Orisa", "Winston", "Roadhog", "D.Va", "Zarya", "Reinhardt", "Wrecking Ball")
supports <- c("Lúcio", "Mercy", "Zenyatta", "Brigitte", "Moira", "Ana", "Sigma", "Baptiste")

dps_records <- df %>% filter(hero %in% dpss)
tank_records <- df %>% filter(hero %in% tanks)
support_records <- df %>% filter(hero %in% supports)
```


For dps players, I will stick to the official measure of a Fleta deadlift: involved in more than 50% percent of team total final blows in a match. 

```{r}
dps_deadlifts <- dps_records %>% 
  filter(stat_name == "Final Blows") %>% 
  mutate(final_blows = stat_amount) %>% 
  select(-stat_name, -stat_amount) %>%
  add_count(date, team, wt = final_blows, name = "team_final_blows") %>%
  filter(team_final_blows > 10) %>% 
  group_by(date, player, team_final_blows) %>%
  summarize(player_final_blows = sum(final_blows)) %>%
  ungroup() %>%
  mutate(rate = player_final_blows / team_final_blows) %>%
  filter(rate > 0.5) 

dps_deadlifts
```



