---
title: Analyzing International Debt Statistics
date: '2020-05-05'
categories: [misc]
description: Analyze international debt from the World Bank using SQL, and make ggplot2 visulizations
last-modified: '2022-09-02'
html-math-method:
  method: plain
---

In this post, I use SQL to retrieve and analyze international debt data collected by The World Bank. The dataset contains information about the amount of debt (in USD `r emo::ji("dollar")`) owed by developing countries across several categories. In fact, I adopted this from one [DataCamp project](https://learn.datacamp.com/projects/754) without following its instructions. The project is still insightful and well-written, though.  Also, the R Markdown documentation has a section on how to [embed SQL chunks](https://bookdown.org/yihui/rmarkdown/language-engines.html#sqls)

## SQL quries

```{r}
#| include: false
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  collapse = TRUE,
  tab.cap = NA
)

knitr::opts_knit$set(sql.max.print = NA)
```

Connect to the postgres database using DBI.

```{r}
library(RPostgreSQL)
library(DBI)

con <- dbConnect(DBI::dbDriver("PostgreSQL"),
  dbname = "qiushi",
  host = "localhost",
  port = 5432,
  user = Sys.getenv("PG_USER"),
  password = Sys.getenv("PG_PASSWORD")
)
```

After successful connection, let's start by `CREATE` the `international_debt` table, and load data into R as well.

```{sql connection='con'}
#| eval: false
CREATE TABLE international_debt (
	country_name varchar(50),
	country_code varchar(10),
	indicator_name varchar(100),
	indicator_code varchar(20),
	debt decimal(12, 1)
)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)

international_debt <- readr::read_csv("https://github.com/qiushiyan/blog-data/blob/main/international_debt.csv?raw=true")
sci_labeller <- label_number(scale_cut = cut_short_scale())
```


Then we can upload debt data into that table. If you happen to be a datacamp subscriber, [here](https://support.datacamp.com/hc/en-us/articles/360020444334-How-to-Download-Project-Datasets) are some instructions on how to download the data. ^[The following chunk is not a real SQL query but plain text. The knitr SQL engine currently only looks for the keywords that are among `INSERT`, `UPDATE`, `DELETE`, `CREATE` and `SELECT`. You have to run the command inside the database.]

```sql
COPY international_debt
FROM 'path/to/csv/file'
WITH (FORMAT csv, header)
```

`international_debt` has debt information about 124 countries and 4714 rows in total, with each row being one type of debt statistics owed by one country or region.

::: {.column-page}

```{sql connection='con'}
-- a glance a debt data
SELECT *
FROM international_debt
LIMIT 10
```

:::

```{sql connection='con'}
-- how many countries
SELECT COUNT(DISTINCT country_code) as n_countries FROM international_debt
```

```{sql connection='con'}
-- how many reords
SELECT COUNT(*) AS n_records FROM international_debt
```



The `indicator_code` column represents the category of these debts. Knowing about these various debt indicators will help us to understand the areas in which a country can possibly be indebted to.

```{sql connection='con'}
SELECT DISTINCT indicator_code, indicator_name FROM international_debt
```

Now, I come to answer questions involving some simple calculations

- What is the total amount of debt of all types? This is a measure of the health of the global economy.

```{sql connection='con'}
SELECT ROUND(SUM(debt), 2) AS total_debt FROM international_debt
```

- Which country has the highest total debt?

```{sql connection='con', tab.cap = "Countries with highest debt"}
SELECT country_name, SUM(debt) AS total_debt
FROM international_debt
GROUP BY country_name
ORDER BY total_debt DESC
LIMIT 20
```

Here we see the top 20 countries with highest overall debt. In fact, some of the entries in `country_name` are not countries but regions, such "South Asia", "Least developed countries: UN classification" and "IDA only".

Now that we know China is in most debt, we could break China's dbet down to see the proportion for which different types of loan accounted.

```{sql connection='con'}
SELECT  indicator_name, debt,
        (debt / sum(debt) OVER()) AS proportion
FROM international_debt
WHERE country_name = 'China'
ORDER BY proportion DESC
```

Two of all categories of debt, long-term and private nonguaranteed principle repayments  on external debt take up more than 50% of China's total debt.

We can dig even further to find out on an average how much debt a country owes. This will give us a better sense of the distribution of the amount of debt across different indicators.

```{sql connection='con'}
SELECT indicator_name, avg(debt) AS mean_debt
FROM international_debt
GROUP BY indicator_name
ORDER BY mean_debt DESC
```

A bit of visualization might help here, I' ll make a density plot of mean debt across all indicators.

```{r}
#| fig-width: 8.5
#| fig-height: 5.5
p <- international_debt %>%
  group_by(indicator_name) %>%
  summarize(mean_debt = mean(debt)) %>%
  ggplot() +
  geom_density(aes(mean_debt), fill = "midnightblue", alpha = 0.4) +
  scale_x_continuous(labels = sci_labeller) +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  labs(
    title = "Distribution of the average debt across different indicators",
    y = NULL,
    x = NULL
  )

p
```

One may notice that principle repayment of long term debts tops the table of average debt and debt proportion of China. As such, we can find the top 10 countries with highest amount of debt in the category of long term debts (`DT.AMT.DLXF.CD`)

```{sql connection='con'}
SELECT DISTINCT country_name
FROM international_debt
WHERE country_name IN (
    SELECT country_name
    FROM international_debt
    WHERE indicator_code = 'DT.AMT.DLXF.CD'
    ORDER BY debt DESC
    LIMIT 10
)
```


We saw that long-term debt is the topmost category when it comes to the average amount of debt. But is it the most common indicator in which the countries owe their debt?


```{sql connection='con'}
SELECT indicator_name, COUNT(indicator_name) As n_indicator
FROM international_debt
GROUP BY indicator_name
ORDER BY n_indicator DESC
```

Turns out it is the second most common category of debt. But what is the average amount of the most common debt type, `DT.INT.OFFT.CD`?

```{sql connection='con'}
SELECT avg(debt) as mean_debt
FROM international_debt
WHERE indicator_code = 'DT.INT.OFFT.CD'
```

```{r}
#| echo: false
#| fig-width: 8.5
#| fig-height: 5.5
df1 <- tibble(
  x = 297677339,
  y = 0.00000000065,
  label = "Average amount of the most common debt category"
)

df2 <- tibble(
  x = 5904868401,
  y = 0.0000000003,
  label = "Average amount of long term debt"
)

p +
  geom_vline(xintercept = 297677339, size = 2, color = "red") +
  geom_vline(xintercept = 5904868401, size = 2, color = "red") +
  ggrepel::geom_text_repel(aes(x = x, y = y, label = label), nudge_x = 1200000000, data = df1) +
  ggrepel::geom_text_repel(aes(x = x, y = y, label = label), data = df2)
```


By inspecting the six indicators in which all the countries listed in our dataset have taken debt (`n_indicator = 124`), we have a clue that all these countries are suffering from some common economic issues. Another problem is what is the most serious issues each country has? We can look into this by retrieving maximum of debt of all categories of each country.

::: {.column-page}


```{sql connection='con'}
-- some countries have tied max debt on multiple categories
WITH max_debt AS (
	SELECT country_name, max(debt) AS maximum
	FROM international_debt
	GROUP BY country_name
	HAVING max(debt) <> 0
)
SELECT  max_debt.country_name, indicator_name, maximum FROM max_debt
  LEFT JOIN (SELECT country_name, indicator_name, debt FROM international_debt) AS debt
	ON max_debt.maximum = debt.debt
	AND max_debt.country_name = debt.country_name
ORDER BY maximum DESC
LIMIT 20
```

:::

## Visualization: countries in most debt

Finally, let's make a plot again to show the top 20 countries with highest debt, as in table 1, plus the specific category in which they take highest debt in.
This time I exclude non-country entries.

```{r}
# prepare data for plot
maximum_category <- international_debt %>%
  group_by(country_name) %>%
  slice_max(order_by = debt, n = 1) %>%
  distinct(country_name, .keep_all = TRUE) %>%
  select(country_name, indicator_name)

countries <- international_debt %>%
  filter(!country_name %in% c(
    "South Asia",
    "Least developed countries: UN classification",
    "IDA only"
  )) %>%
  group_by(country_name) %>%
  summarize(total_debt = sum(debt)) %>%
  slice_max(order_by = total_debt, n = 20) %>%
  left_join(maximum_category)

countries
```

```{r}
# plotting
library(ggchicklet)
library(ggtext)
library(showtext)

# load fonts
font_add_google("Overpass Mono", "Overpass Mono")
font_add_google("Roboto Condensed", "Roboto Condensed")
showtext_auto()
```

```{r}
#| column: page
#| fig-cap: highest contributions from long term <span style='color:#F0D8C0'>repayments</span> or <span style='color:#6078A8'>disbursements</span>
#| fig-width: 14
#| fig-height: 11

ggplot(countries) +
  geom_chicklet(
    aes(
      x = forcats::fct_reorder(country_name, total_debt),
      y = total_debt,
      fill = indicator_name
    ),
    color = NA, width = 0.8
  ) +
  geom_text(
    aes(
      country_name, total_debt,
      label = sci_labeller(total_debt)),
      color = "white", nudge_y = -10000000000, family = "Overpass Mono"
  ) +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale(), prefix = "$")) +
  hrbrthemes::theme_modern_rc() +
  nord::scale_fill_nord(palette = "afternoon_prarie", name = NA) +
  coord_flip(clip = "off") +
  labs(
    x = NULL,
    y = NULL,
    title = "Top 20 Countries with Highest Total Debts"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 28, family = "Roboto Condensed"),
    plot.title.position = "plot",
    plot.subtitle = element_markdown(family = "Roboto Condensed"),
    axis.text.x = element_text(face = "bold", size = 14),
    axis.text.y = element_text(face = "bold", size = 18),
    panel.grid.major.y = element_blank()
  )
```
